The Design, Semantics, and Implementation of the Synchronous Language \CEU

### 2016-03-26 paper.tex
...

@@

\begin{comment}
Basically, they map

Basically, it is composed by instructions to define a pending trail (\code{tkins}),
a pending event(using \code{set\_c} to set a value in the memory),
a check operation to verify the \code{par/and} termination(\code{chkret}),
a pending timer event (\code{clken}), and some others like to clear trails data or suspend execution.
All pending entries are managed by the scheduler, that also responds for timer and input events.
All trails are executed as a TinyOS task, this ensure that each trail runs to completion avoiding race condition between then.

During compilation process the script is broken in trails in the same way it is 
done in the \CEU compiler explained in Section 4.
The Terra virtual machine implements a similar \CEU scheduler to execute the 
script trails, but the VM
uses the real trails address instead of labels. In this way is possible to jump directly to each script
partitions without using a dispatcher.

Figure~\ref{lst.impl.assembly} presents the bytecode generated for the initial 
part for the example of Figure~\ref{lst.impl.trails}.
After initializing the trail and event slots in the memory, the VM engine starts the execution from the \code{main} label.
In this example it registers in the trails slots the label addresses for the two block of the \code{par/and} control (lines 3--6).
The scheduler, first, executes the code in the label \code{and\_sub\_1} that registers the label \code{awk\_e} into the slot of event \code{e} (lines 8--10).
Then, executes the code in the label \code{and\_sub\_2} that registers the label \code{awk\_temp} into the slot of the event \code{I\_TEMP} (lines 17--18).
In this moment the VM engine suspends its execution waiting for some new input or timer event.
In the case of this example, the engine waits for the \code{I\_TEMP} event to continue the script execution.

\begin{figure}
\begin{minipage}[t]{0.91\linewidth}
\begin{lstlisting}[language=C,numbers=left,xleftmargin=2.5em,keywords={tkins,end,set_c,ushort,ubyte,exec}]
  <...> // Initializes the trails and events slots

:main -- Activate TRAILs 0 & 1
  tkins and_sub_1           // Activate TRAIL 0 - lbl and_sub_1
  tkins and_sub_2           // Activate TRAIL 1 - lbl and_sub_2
  end                       // Suspend execution

:and_sub_1
  set_c ushort slot_e awk_e // await e
  end                       // Suspend execution

:awk_e
  set_c ubyte slot_1 1      // Close TRAIL 0 - lbl and_sub_1
  exec chkret               // goto check par/and termination

:and_sub_2
  set_c ushort I_TEMP awk_temp // await TEMP
  end                       // Suspend execution

  <....> // Other entries: chkret, awk_TEMP, etc..
\end{lstlisting}
\end{minipage}
\caption{
%{\small %\textmd{
Assembly code for a \code{par/and} control structure.
%}%}
\label{lst.impl.assembly}
}
\end{figure}

Instead of generating code in \C, Terra generates the vmt assembly code using 
the original \CEU labels for trails entry points.
In the final phase for code generation, Terra compiler generates the vmt 
bytecode file using the program address in place of \CEU labels.
The use of address allows the \CEU scheduler to execute directly a trail 
without a index table to replace the switch-case.
Related to memory layout and trail allocation, Terra uses the same data 
organization approach for all \CEU controls and variables.
All \CEU data and control structures for a specific program are determined at 
compile time.
Terra arranges all these structures in the beginning of virtual machine memory 
space.
The original \CEU run time control functions are maintained with a small 
modification where accesses to control structures have an off-set relative to 
the begin of the virtual machine memory space.
\end{comment}

@@

\begin{comment}
- Kiel Esterel compiler

- We are not concerned with WCET:
Predictable system design is concerned with the challenge o
f building sys-
tems in such a way that requirements can be guaranteed from th
e design. This
means that an off-line analysis should demonstrate satisfa
ction of timing require-
ments, subject to assumptions made on operating conditions
foreseen for the sys-
tem [Stankovic and Ramamritham 1990]

http://embedded.cs.uni-saarland.de/publications/BuildingTimingPredictableEmbeddedSystems.pdf
survey of timeing-predictable embedded systems,
synchronous languages do some at the lang semantic level
\cite{pret.building}
\end{comment}

>>>>

Single-thread dispatching may not be suitable for hard real-time activities.
In a previous work~\cite{ceu.sensys13}, we measure how synchronous lengthy 
computations in C (e.g., hashing and compression) can block the scheduler and 
affect higher-priority activities such as a radio driver.
In such cases, the system requires careful testing to avoid undersized hardware 
deployment.

@@

@@
